Hydrodynamics and rheology of a vesicle doublet suspension

### 2019-07-20 vesicle_code/docs/adhesion/adhesion.tex
edits to adhesion.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Adhesion Model}
%
\begin{figure}[htp]
  \begin{tabular}{@{}p{0.3\linewidth}@{\qquad}p{0.3\linewidth}@{\qquad}p{0.3\linewidth}@{}}
  \subfigimg[height=\linewidth]{(a)}{figs/adhesionPotential.pdf} &
  \subfigimg[height=\linewidth]{(b)}{figs/configCartoon.pdf} &
  \subfigimg[height=\linewidth]{(c)}{figs/Adhesion_Force.png}
  \end{tabular}
  \caption{\label{fig:adhesionModel} (a) The adhesion potential
  $\phi(z)$ in equation~\eqref{eq:adhesion_potential}.  Distances less
  than $\delta$ are repulsive and distances greater than $\delta$ are
  attractive.  (b) Each point on a vesicle is repelled or attracted by
  all points on the other vesicles.  (c) The resulting adhesive force is
  found by integrating over all points on all other vesicles
  (equation~\eqref{eqn:adhesionForce}).  The color is the dot product of
  the adhesion force with the vector joining the vesicles.  Therefore,
  the green and yellow regions are repelled by the other vesicle while
  the blue regions are attracted to the other vesicle.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a quiescent flow} 
\label{sec:qflow} 
\todo[inline]{Bryan is here}
We consider two identical vesicles suspended in a quiescent flow.
Without any external forcing such as an imposed electric field or a
fluid flow in the far-field, the long-range attraction pulls the
vesicles together until their separation distance is close to $\delta$.
The L.-J.~type potential prohibits physical contact between the two
membranes and instead keeps them close to the separation distance
$\delta$.  In Section~\ref{subsec:qflow_draining_times}, we compute the
expected time for the vesicles to reach an equilibrium configuration and
compare the analysis with numerical simulations.  In
Section~\ref{subsec:qflow_adhesion_parameters}, we characterize the
effects of the adhesion parameters and the vesicles' reduced area on the
shape of the contact region and the bending and adhesive energies.

@@

When the force on each vesicle is a function of the separation distance
$h$ of the form
\begin{align}
  \label{eq:Fh}
  F(h) \sim \mathcal{H} \frac{m\delta^m}{h^{m+1}}
    \left(1-\left(\frac{\delta}{h}\right)^{n-m}\right),
\end{align}
with integers $m>n\ge 2$.  Attraction is dominant at ``large"
distances ($h > \delta$) while repulsion is dominant at ``small"
distances ($h<\delta$).  Integration of equation~\eqref{eq:dhdt} with
the $F(h)$ as defined in equation~\eqref{eq:Fh} gives the relationship
between $t$ and $h$.  In this case, the relationship involves an
integral of a function of the dimensionless variable $\bar{h} \equiv
h/\delta$:
\begin{align*}
  t\frac{K_a^{2/3}}{\mu R_0^{10/3}}& = 
    \int^{h(t)}_{h(0)}-\frac{dh}{h^3 F^{1/3}} = 
    -\frac{1}{m^{1/3}\mathcal{H}^{1/3}\delta^{5/3}}\int^{\bar{h}(t)}_{\bar{h}(0)}
      \frac{d \bar{h}}{\bar{h}^{(8-m)/3}(\bar{h}^{m-n}-1)^{1/3}}.
\end{align*}
Solving for $t$,
\begin{align}
\label{eq:teq_scaling}
  t\sim \frac{\mu R_0^{10/3}}{K_a^{2/3}}\frac{1}{ m^{1/3} \mathcal{H}^{1/3}\delta^{5/3}},
\end{align}
and equation~\eqref{eq:teq_scaling} tells us that both the adhesion
strength $\mathcal{H}$ and the separation distance $\delta$ affect the
time it takes for a pair of vesicles to reach the separation distance
$\delta$ under adhesion force $F$ in equation~\eqref{eq:Fh}.  In
particular, the duration $t$ is proportional to (i) the separation
distance $\delta^{-5/3}$ (thus $t\rightarrow \infty$ as
$\delta\rightarrow 0$), and (ii) the adhesion strength
$\mathcal{H}^{-1/3}$ (as in the constant forcing
case~\cite{RamachandranLeal2010_PoF}).  From the above analysis we also
expect that the scaling of $t$ with respect to $\delta$ depends on the
adhesion potential, while the scaling with respect to adhesion strength
is independent of the specific form of the potential.

@@

In this work we focus on weakly adhesive vesicles whose equilibrium
shapes in a quiescent flow are shown in the left plot of
Figure~\ref{fig:Dec18_vesicle_shape} for three reduced areas: $\Delta
A=0.95$, $0.75$, and $0.6$ with $\mathcal{H}=5$ and $\delta = 0.2$. 
\begin{figure}
%  \includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_vesicle_shape_vs_rA_00.jpeg}
%  \includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_vesicle_shape_vs_rA_01.jpeg}
   \includegraphics[keepaspectratio=true,scale=0.45]{figs/Dec18_vesicle_shape_vs_rA_composite.png}
  \caption{\label{fig:Dec18_vesicle_shape} The equilibrium
  configurations of a doublet of identical vesicles under adhesion in a
  quiescent flow at three values of $\Delta A$.  The vesicle reduced
  areas of $\Delta A=0.95$, $0.75$, and $0.6$ are labeled.  The vesicle
  length is fixed at $2\pi$, the dimensionless bending modulus
  $\kappa_b=1$, the Hamaker constant $\mathcal{H}=5$, and the
  separation distance $\delta=0.2$.  The color coding is the tension
  along the vesicle.  The overall equilibrium shapes of two vesicles
  under adhesive interactions vary with the reduced area.  The right
  figure shows the  membrane shapes in the contact region. The color
  coding is the adhesive force projected along the center-of-mass
  vector.} 
\end{figure}
The equilibrium vesicle shape for $\Delta A=0.95$ is the circular cap
with a flat contact region.  This is similar to the observed shapes of
two vesicles under strong adhesive interaction
in~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.  When
vesicles are more deflated with a reduced area $\Delta A = 0.75$, the
equilibrium vesicle shape is elongated with a bigger contact region.
This is consistent with the equilibrium shapes of a vesicle doublet
under a $(m,n) = (12,6)$
L.-J.~potential~\cite{FlormannAouane2017_SciReports}.  As the reduced
area decreases further, we observe undulation of the vesicle membrane on
the non-contact side while the contact region remains flat.  The color
coding along each curve is the tension of the vesicle membrane.  We
observe that the membrane tension in the contact region is very
negative, indicating a dominant compression of membrane when the
adhesion force is strong to keep the vesicles bound together.

The right plot of Figure~\ref{fig:Dec18_vesicle_shape} is a zoom (not
to-scale) of the membranes in the contact region, where the membranes
are not perfectly flat at all three values of reduced area: We observe
that the membranes are slightly curved with a dip at the edge.  Such
membrane undulation in the contact region is predicted by lubrication
analyses on an elastic membrane under adhesion with a solid
substrate~\cite{BlountMiksisDavis2013_PRSa, YoungStone2017_PRF}.  
%
Results from the lubrication analysis show that
this dip and slightly curved shape in the contact region are independent
of the adhesion strength. The high membrane curvature at the edge may
pose a problem for using the contact angle there to estimate the
adhesion strength. This inspires us to investigate the possibility of
using a dynamic fluid trap to measure the adhesion strength (see
Section~\ref{sec:eflow}).

Also the dip at the edge of contact region is
related to (but different from) the buckling of membrane under strong
adhesion: A closer inspection on the dip at the edge shows that the
membrane distance is smaller than the neutral separation distance
$\delta$ there for more deflated vesicles.
For $\Delta A =0.6$ and $0.75$  the adhesion force is mostly attractive
 in the contact region except at the edge, where the adhesive
force turns repulsive.  For $\Delta A=0.95$, however, the adhesion force is mostly repulsive,
leading to large tension in the rest of the vesicle membrane.

\begin{figure}
  \begin{tabular}{@{}p{0.45\linewidth}@{\quad}p{0.45\linewidth}@{}}
  \subfigimg[width=\linewidth]{(a)}{figs/Dec18_Ebleft_Eadhright_vs_rA_adR0p2_adS502.jpeg} &
  \subfigimg[width=\linewidth]{(b)}{figs/Dec18_EbEadh_vs_rA_adR0p2_adS502.jpeg}
  \end{tabular}
%\includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_Ebleft_Eadhright_vs_rA_adR0p2_adS502.jpeg}
%\includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_EbEadh_vs_rA_adR0p2_adS502.jpeg}
  \caption{\label{fig:Dec18_vesicle_equilibrium1}   (a) The total
  bending energy (solid curve) and adhesion energy (dash-dotted curve)
  plotted against the reduced area $\Delta A$. (b) The sum of two
  energies plotted against reduced area $\Delta A$. The two vesicles in
  the doublet are of the same length and reduced area.  The Hamaker
  constant is $\mathcal{H} = 5$ and the separation distance is $\delta =
  0.2$.}
\end{figure}

Figure~\ref{fig:Dec18_vesicle_equilibrium1}(a) shows the total bending
energy and adhesion energy as a function of the reduced area. The
smaller the reduced area, the more vesicle area is available for
deformation and thus the bending energy is higher.  In contrast, the
total adhesion energy becomes more negative as the reduced area
decreases.  The sum of two energies is plotted in
Figure~\ref{fig:Dec18_vesicle_equilibrium1}(b), where a local minimum in
the total energy is found around $\Delta A = 0.85$.

@@

  \includegraphics[height = 0.2\textwidth,trim={1cm 0cm 1cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_tension_image01.png}
  \includegraphics[height = 0.2\textwidth,trim={4cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_tension_image02.png}
  \includegraphics[height = 0.2\textwidth,trim={4cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_tension_image03.png}
  \includegraphics[height = 0.2\textwidth,trim={4cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_tension_image04.png}
  \includegraphics[height = 0.2\textwidth,trim={3.5cm 0cm 0cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_tension_image05.png}
  \caption{\label{fig:extensionalNonsymmetric}Dynamics of a vesicle pair
  in a planar extensional flow. The Hamaker constant is ${\cal H} = 0.7$
  for the first three rows, and ${\cal H}=0.6$ for the bottom row. From
    top to bottom, the initial vertical displacement is $\Delta y=0$,
    $0.02$, $0.1$, and $0.1$. The reduced area is $\Delta A = 0.75$, the
    extension rate $\dot\epsilon= 0.09$, and the adhesion distance
    $\delta = 0.4$.  The color coding is the tension.}
  \end{figure}

The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous
drop~\cite{KantslerSegreSteinberg2007_PRL,KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}.  Here we focus on
a planar extensional flow  where the fluid velocity field
$\uu=\dot\epsilon(-x,y)$ is centered at the origin ($(x,y)=(0,0)$), the
stagnation point where the fluid flow converges horizontally and
diverges vertically.  Placed symmetrically at the stagnation point of an
extensional flow, a single vesicle may remain steady and symmetric for
low extension rate $\dot\epsilon$.  For sufficiently large extension
rate $\dot\epsilon$ the vesicle may undergo asymmetric
deformation~\cite{KantslerSegreSteinberg2008_PRL,Narsimhan2014_JFM,DahlNarsimhanGouveia2016_SoftMatt},
similar to a surfactant-laden viscous drop under a planar extensional
flow~\cite{JanssenBoonAgterof1997_AIChE,HuPineLeal2000_PoF}.  To the
best of our knowledge, no studies on the interactions between two
vesicles in an extensional flow are available in the literature.  In the
following we first illustrate how two vesicles interact with each other
in a planar extensional flow. We then illustrate how the adhesion
strength between membranes can be measured non-intrusively by using the
fluid trap.

At the beginning of the first set of simulations (left panels in
Figure~\ref{fig:extensionalNonsymmetric}) we place a vesicle at
$(x,y)=(0,0)$, and the other vesicle at $(x,y) = (3,\Delta y)$ where the
initial vertical displacement $\Delta y$ varies from $\Delta y=0$ for
the top row, $\Delta y = 0.02$ for the second row, to $\Delta y =0.1$
for the bottom two rows in Figure~\ref{fig:extensionalNonsymmetric}.
The Hamaker constant is ${\cal H} = 0.7$ for the top three rows, and is
${\cal H} =0.6$ for the bottom row.  The color coding is the membrane
tension with the color bar on the right of each row. We note that the
negative tension in an extensional flow is consistent with the destabilizing tension 
for a vesicle with finite bending forces~\cite{Narsimhan2014_JFM}.
%and the speculation for a wrinkled membrane in the elongational flow~\cite{KantslerSegreSteinberg2007_PRL}.

@@

For $\Delta y>0$ we observe similar vesicle dynamics only at the
beginning when the right vesicle moves towards the stagnation point and
the left vesicle is pushed to the left of the stagnation point.  As both
vesicles come to close vicinity of the stagnation point, the elevated
right vesicle ($\Delta y > 0$) is pulled up by the diverging flow in the
$y$ direction, causing rotation of both vesicles.  We also observe that
the larger the initial vertical displacement, the larger the rotation
and the faster both vesicles move away from the stagnation point. For
sufficiently strong adhesion (${\cal H}=0.7$ for the first three rows in
Figure~\ref{fig:extensionalNonsymmetric}), the two vesicles rotate and
move away from the stagnation point as a doublet until the end of
simulations ($t=100$).  As the two vesicles move farther away from the
stagnation point, the extensional flow becomes more effective in pulling
the two drops apart as shown in the third row of
Figure~\ref{fig:extensionalNonsymmetric}.  For ${\cal H}=0.6$ (bottom
row in Figure~\ref{fig:extensionalNonsymmetric})  the two vesicles are
separated by the diverging flow as early as $t=75$, and are clearly not
a doublet at the end of simulation.

Here we provide two main observations drawn from the results in
Figure~\ref{fig:extensionalNonsymmetric}: (1) The membrane tension is
highly non-uniform on the vesicles, leading to a significant Marangoni
stress along the membrane as in the case of a surfactant-laden drop at
the center of an extensional flow. Thus it is reasonable to expect that
the flow around the vesicles is significantly altered.  (2) In an
extensional flow, vesicles are prone to move away from the stagnation
point.  A small initial vertical displacement $\Delta y$ in one  of the
vesicles causes both vesicles to not only rotate but also move away from
the center. Once vesicles move off the horizontal axis, their
hydrodynamics is dominated by the diverging flow as they move farther
away.  Thus in the microfluidic experiment it is often more desirable to
trap the vesicle around the center of an extensional
flow~\cite{Spjut2010_MSThesis_Chapter3}. With an active control
algorithm to place the stagnation point at a desirable location by
adjusting the flow with a feedback
loop~\cite{BentleyLeal1986_JFMa,Johnson-Chavarria2011_EMJ}, a particle
can be trapped at the stagnation point for long time scales to
facilitate image acquisition or other detailed measures such as particle
image velocimetry of flow inside and around the particle.

@@

We consider vesicle doublets with reduced areas 0.70, 0.75, 0.80, 0.85,
0.90, and 0.95, all with a length of $2\pi$, and we vary the extension
rate ($\dot\epsilon$) between $10^{-2}$ and $10^{-1}$.  Since the
stagnation point can be controlled in an experimental
setting~\cite{BentleyLeal1986_JFMa,Johnson-Chavarria2011_EMJ}, we mimic
the active control of the microfluidic experiments by moving the center
of the doublet at each time step so that the stagnation point occurs
exactly in the middle of the doublet.  With this adjustment, the
vesicles either remain as a doublet in the fluid trap centered around
the stagnation point, or the doublet is broken and the vesicles separate
from one another.

@@

\begin{figure}
  \begin{tabular}{@{}p{0.45\linewidth}@{\quad}p{0.45\linewidth}@{}}
  \subfigimg[width=\linewidth]{(a)}{figs/shear_adR1em1_ra090_phaseDiagram.pdf} &
  \subfigimg[width=\linewidth]{(b)}{figs/shear_adR4em1_ra090_phaseDiagram.pdf}
  \end{tabular}
  \caption{\label{fig:sflow_phase_diagram} Phase diagram of vesicle
  hydrodynamics in a shear flow with $\delta = 0.1$ (a) and $\delta =
  0.4$ (b) with reduced area $\Delta A = 0.9$. The critical Hamaker
  constant $\mathcal{H}$ for binding of two vesicles in a shear flow
  depends on the shear rate $\dot\gamma$.  For a doublet to form, lower
  separation distances require a stronger Hamaker constants at a given
  shear rate.  The stability limit appears to be linear, indicating that
  there is potentially a stability criteria that depends on a
  dimensionless number involving the ratio of the Hamaker constant and
  shear rate.}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
  \begin{tabular}{@{}p{0.3\linewidth}@{\qquad}p{0.3\linewidth}@{\qquad}p{0.3\linewidth}@{}}
  \subfigimg[height=\linewidth]{(a)}{figs/adhesionPotential.pdf} &
  \subfigimg[height=\linewidth]{(b)}{figs/configCartoon.pdf} &
  \subfigimg[height=\linewidth]{(c)}{figs/Adhesion_Force.png}
  \end{tabular}
  \caption{\label{fig:adhesionModel} (a) The adhesion potential
  $\phi(z)$ in equation~\eqref{eq:adhesion_potential}.  Distances less
  than $\delta$ are repulsive and distances greater than $\delta$ are
  attractive.  (b) Each point on a vesicle is repelled or attracted by
  all points on the other vesicles.  (c) The resulting adhesive force is
  found by integrating over all points on all other vesicles
  (equation~\eqref{eqn:adhesionForce}).  The color is the dot product of
  the adhesion force with the vector joining the vesicles.  Therefore,
  the green and yellow regions are repelled by the other vesicle while
  the blue regions are attracted to the other vesicle.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a quiescent flow} 
\label{sec:qflow} 
We consider two identical vesicles suspended in a quiescent flow.
Without any external forcing such as an imposed electric field or a
fluid flow in the far-field, the long-range attraction pulls the
vesicles together until their separation distance is close to $\delta$.
The L.-J.~type potential prohibits physical contact between the two
membranes and instead keeps them close to the separation distance
$\delta$.  In Section~\ref{subsec:qflow_draining_times}, we compute the
expected time for the vesicles to reach an equilibrium configuration and
compare the analysis with numerical simulations.  In
Section~\ref{subsec:qflow_adhesion_parameters}, we characterize the
effects of the adhesion parameters and the vesicles' reduced area on the
shape of the contact region and the bending and adhesive energies.

@@

\todo[inline]{This first sentence is incomplete}
When the force on each vesicle is a function of the separation distance
$h$ of the form
\begin{align}
  \label{eq:Fh}
  F(h) \sim \mathcal{H} \frac{m\delta^m}{h^{m+1}}
    \left(1-\left(\frac{\delta}{h}\right)^{n-m}\right),
\end{align}
with integers $m>n\ge 2$.  Attraction is dominant at ``large"
distances ($h > \delta$) while repulsion is dominant at ``small"
distances ($h<\delta$).  Integration of equation~\eqref{eq:dhdt} with
the $F(h)$ as defined in equation~\eqref{eq:Fh} gives the relationship
between $t$ and $h$.  In this case, the relationship involves an
integral of a function of the dimensionless variable $\bar{h} \equiv
h/\delta$:
\begin{align*}
  t\frac{K_a^{2/3}}{\mu R_0^{10/3}}& = 
    \int^{h(t)}_{h(0)}-\frac{dh}{h^3 F^{1/3}} = 
    -\frac{1}{m^{1/3}\mathcal{H}^{1/3}\delta^{5/3}}\int^{\bar{h}(t)}_{\bar{h}(0)}
      \frac{d \bar{h}}{\bar{h}^{(8-m)/3}(\bar{h}^{m-n}-1)^{1/3}}.
\end{align*}
Solving for $t$,
\begin{align}
\label{eq:teq_scaling}
  t\sim \frac{\mu R_0^{10/3}}{K_a^{2/3}}\frac{1}{ m^{1/3} \mathcal{H}^{1/3}\delta^{5/3}},
\end{align}
and equation~\eqref{eq:teq_scaling} tells us that both the adhesion
strength $\mathcal{H}$ and the separation distance $\delta$ affect the
time it takes for a pair of vesicles to reach the separation distance
$\delta$ under adhesion force $F$ in equation~\eqref{eq:Fh}.  In
particular, the duration $t$ is proportional to (i) the separation
distance $\delta^{-5/3}$ (thus $t\rightarrow \infty$ as
$\delta\rightarrow 0$), and (ii) the adhesion strength
$\mathcal{H}^{-1/3}$ (as in the constant forcing
case~\cite{RamachandranLeal2010_PoF}).  From the above analysis we also
expect that the scaling of $t$ with respect to $\delta$ depends on the
adhesion potential, while the scaling with respect to adhesion strength
is independent of the specific form of the potential.

@@

In this work we focus on weakly adhesive vesicles whose equilibrium
shapes in a quiescent flow are shown in the left plot of
Figure~\ref{fig:Dec18_vesicle_shape} for three reduced areas: $\Delta
A=0.95$, $0.75$, and $0.6$ with $\mathcal{H}=5$ and $\delta = 0.2$. 
\begin{figure}
%  \includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_vesicle_shape_vs_rA_00.jpeg}
%  \includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_vesicle_shape_vs_rA_01.jpeg}
   \includegraphics[keepaspectratio=true,scale=0.45]{figs/Dec18_vesicle_shape_vs_rA_composite.png}
  \caption{\label{fig:Dec18_vesicle_shape} The equilibrium
  configurations of a doublet of identical vesicles under adhesion in a
  quiescent flow at three values of $\Delta A$.  The vesicle reduced
  areas of $\Delta A=0.95$, $0.75$, and $0.6$ are labeled.  The vesicle
  length is fixed at $2\pi$, the dimensionless bending modulus is
  $\kappa_b=1$, the Hamaker constant is $\mathcal{H}=5$, and the
  separation distance is $\delta=0.2$.  The color coding is the tension
  along the vesicle.  The overall equilibrium shapes of two vesicles
  under adhesive interactions vary with the reduced area.  The right
  figure shows the  membrane shapes in the contact region. The color
  coding is the adhesive force projected along the center-of-mass
  vector.} 
\end{figure}
The equilibrium vesicle shape for $\Delta A=0.95$ is the circular cap
with a flat contact region.  This is similar to the observed shapes of
two vesicles under strong adhesive interaction
in~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.  When
vesicles are more deflated with a reduced area $\Delta A = 0.75$, the
equilibrium vesicle shape is elongated with a bigger contact region.
This is consistent with the equilibrium shapes of a vesicle doublet
under a $(m,n) = (12,6)$
L.-J.~potential~\cite{FlormannAouane2017_SciReports}.  As the reduced
area decreases further, we observe undulation of the vesicle membrane on
the non-contact side while the contact region remains flat.  The color
coding along each curve is the tension of the vesicle membrane.  We
observe that the membrane tension in the contact region is very
negative, indicating a dominant compression of membrane when the
adhesion force is strong to keep the vesicles bound together.

The right plot of Figure~\ref{fig:Dec18_vesicle_shape} is a zoom (not
to-scale) of the membranes in the contact region, where the membranes
are not perfectly flat at all three values of reduced area. We observe
that the membranes are slightly curved with a dip at the edge.  Such
membrane undulation in the contact region is predicted by lubrication
analyses on an elastic membrane under adhesion with a solid
substrate~\cite{BlountMiksisDavis2013_PRSa, YoungStone2017_PRF}. Results
from the lubrication analysis show that this dip and slightly curved
shape in the contact region are independent of the adhesion strength.
The high membrane curvature at the edge may pose a problem for using the
contact angle there to estimate the adhesion strength. This inspires us
to investigate the possibility of using a dynamic fluid trap to measure
the adhesion strength (see Section~\ref{sec:eflow}).

Also the dip at the edge of the contact region is related to (but
different from) the buckling of membrane under strong adhesion: A closer
inspection on the dip at the edge shows that the membrane distance is
smaller than the neutral separation distance $\delta$ there for more
deflated vesicles.  For $\Delta A =0.6$ and $0.75$  the adhesion force
is mostly attractive in the contact region except at the edge, where the
adhesive force turns repulsive.  For $\Delta A=0.95$, however, the
adhesion force is mostly repulsive, leading to large tension in the rest
of the vesicle membrane.

\begin{figure}
  \begin{tabular}{@{}p{0.45\linewidth}@{\quad}p{0.45\linewidth}@{}}
  \subfigimg[width=\linewidth]{(a)}{figs/Dec18_Ebleft_Eadhright_vs_rA_adR0p2_adS502.jpeg} &
  \subfigimg[width=\linewidth]{(b)}{figs/Dec18_EbEadh_vs_rA_adR0p2_adS502.jpeg}
  \end{tabular}
%\includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_Ebleft_Eadhright_vs_rA_adR0p2_adS502.jpeg}
%\includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_EbEadh_vs_rA_adR0p2_adS502.jpeg}
  \caption{\label{fig:Dec18_vesicle_equilibrium1} (a) The total bending
  energy (solid curve) and adhesion energy (dash-dotted curve) plotted
  against the reduced area $\Delta A$. (b) The sum of the two energies
  plotted against reduced area $\Delta A$. The two vesicles in the
  doublet are of the same length and reduced area.  The Hamaker constant
  is $\mathcal{H} = 5$ and the separation distance is $\delta = 0.2$.}
\end{figure}

Figure~\ref{fig:Dec18_vesicle_equilibrium1}(a) shows the total bending
energy and adhesion energy as a function of the reduced area. The
smaller the reduced area, the more vesicle area is available for
deformation and thus the bending energy is higher.  In contrast, the
total adhesion energy becomes more negative as the reduced area
decreases.  The sum of the two energies is plotted in
Figure~\ref{fig:Dec18_vesicle_equilibrium1}(b), where a local minimum in
the total energy is found around $\Delta A = 0.85$.

@@

  \includegraphics[height = 0.2\textwidth,trim={1cm 0cm 1cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_tension_image01.png}
  \includegraphics[height = 0.2\textwidth,trim={4cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_tension_image02.png}
  \includegraphics[height = 0.2\textwidth,trim={4cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_tension_image03.png}
  \includegraphics[height = 0.2\textwidth,trim={4cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_tension_image04.png}
  \includegraphics[height = 0.2\textwidth,trim={3.5cm 0cm 0cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_tension_image05.png}
  \caption{\label{fig:extensionalNonsymmetric} The dynamics of a vesicle
  pair in a planar extensional flow. The Hamaker constant is ${\cal H} =
  0.7$ for the first three rows, and ${\cal H}=0.6$ for the bottom row.
  From top to bottom, the initial vertical displacement is $\Delta y=0$,
  $0.02$, $0.1$, and $0.1$. The reduced area is $\Delta A = 0.75$, the
  extension rate $\dot\epsilon= 0.09$, and the adhesion distance $\delta
  = 0.4$.  The color coding is the tension.}
  \end{figure}

The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous
drop~\cite{KantslerSegreSteinberg2007_PRL,
KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM, Narsimhan2014_JFM,
DahlNarsimhanGouveia2016_SoftMatt}.  Here we focus on a planar
extensional flow  where the fluid velocity field
$\uu=\dot\epsilon(-x,y)$ is centered at the origin ($(x,y)=(0,0)$), the
stagnation point where the fluid flow converges horizontally and
diverges vertically.  Placed symmetrically at the stagnation point of an
extensional flow, a single vesicle may remain steady and symmetric for
low extension rate $\dot\epsilon$.  For sufficiently large extension
rate $\dot\epsilon$ the vesicle may undergo asymmetric
deformation~\cite{KantslerSegreSteinberg2008_PRL, Narsimhan2014_JFM,
DahlNarsimhanGouveia2016_SoftMatt}, similar to a surfactant-laden
viscous drop under a planar extensional
flow~\cite{JanssenBoonAgterof1997_AIChE, HuPineLeal2000_PoF}.  To the
best of our knowledge, no studies on the interactions between two
vesicles in an extensional flow are available in the literature.  In the
following we first illustrate how two vesicles interact with each other
in a planar extensional flow. We then illustrate how the adhesion
strength between membranes can be measured non-intrusively by using the
fluid trap.

At the beginning of the first set of simulations (left panels in
Figure~\ref{fig:extensionalNonsymmetric}) we place a vesicle at
$(x,y)=(0,0)$, and the other vesicle at $(x,y) = (3,\Delta y)$ where the
initial vertical displacement $\Delta y$ varies from $\Delta y=0$ for
the top row, $\Delta y = 0.02$ for the second row, to $\Delta y =0.1$
for the bottom two rows in Figure~\ref{fig:extensionalNonsymmetric}.
The Hamaker constant is ${\cal H} = 0.7$ for the top three rows, and is
${\cal H} =0.6$ for the bottom row.  The color coding is the membrane
tension with the color bar on the right of each row. We note that the
negative tension in an extensional flow is consistent with the
destabilizing tension for a vesicle with finite bending
forces~\cite{Narsimhan2014_JFM}.
%and the speculation for a wrinkled membrane in the elongational
%flow~\cite{KantslerSegreSteinberg2007_PRL}.

@@

For $\Delta y>0$ we observe similar vesicle dynamics only at the
beginning when the right vesicle moves towards the stagnation point and
the left vesicle is pushed to the left of the stagnation point.  As both
vesicles come to close vicinity of the stagnation point, the elevated
right vesicle ($\Delta y > 0$) is pulled up by the diverging flow in the
$y$ direction, causing rotation of both vesicles.  We also observe that
the larger the initial vertical displacement, the larger the rotation
and the faster both vesicles move away from the stagnation point. For
sufficiently strong adhesion (${\cal H}=0.7$ for the first three rows in
Figure~\ref{fig:extensionalNonsymmetric}), the two vesicles rotate and
move away from the stagnation point as a doublet until the end of
simulations ($t=100$).  As the two vesicles move farther away from the
stagnation point, the extensional flow becomes more effective in pulling
the two drops apart as shown in the third row of
Figure~\ref{fig:extensionalNonsymmetric}.  For ${\cal H}=0.6$ (bottom
row in Figure~\ref{fig:extensionalNonsymmetric}), the two vesicles are
separated by the diverging flow as early as $t=75$, and are clearly not
a doublet at the end of simulation.

Here we provide two main observations drawn from the results in
Figure~\ref{fig:extensionalNonsymmetric}: (1) The membrane tension is
highly non-uniform on the vesicles, leading to a significant Marangoni
stress along the membrane as in the case of a surfactant-laden drop at
the center of an extensional flow. Thus it is reasonable to expect that
the flow around the vesicles is significantly altered.  (2) In an
extensional flow, vesicles are prone to move away from the stagnation
point.  A small initial vertical displacement $\Delta y$ in one  of the
vesicles causes both vesicles to not only rotate but also move away from
the center. Once vesicles move off the horizontal axis, their
hydrodynamics is dominated by the diverging flow as they move farther
away.  Thus in the microfluidic experiment it is often more desirable to
trap the vesicle around the center of an extensional
flow~\cite{Spjut2010_MSThesis_Chapter3}. With an active control
algorithm to place the stagnation point at a desirable location by
adjusting the flow with a feedback loop~\cite{BentleyLeal1986_JFMa,
Johnson-Chavarria2011_EMJ}, a particle can be trapped at the stagnation
point for long time scales to facilitate image acquisition or other
detailed measures such as particle image velocimetry of flow inside and
around the particle.

@@

We consider vesicle doublets with reduced areas 0.70, 0.75, 0.80, 0.85,
0.90, and 0.95, all with a length of $2\pi$, and we vary the extension
rate ($\dot\epsilon$) between $10^{-2}$ and $10^{-1}$.  Since the
stagnation point can be controlled in an experimental
setting~\cite{BentleyLeal1986_JFMa, Johnson-Chavarria2011_EMJ}, we mimic
the active control of the microfluidic experiments by moving the center
of the doublet at each time step so that the stagnation point occurs
exactly in the middle of the doublet.  With this adjustment, the
vesicles either remain as a doublet in the fluid trap centered around
the stagnation point, or the doublet is broken and the vesicles separate
from one another.

@@

\begin{figure}
  \begin{tabular}{@{}p{0.45\linewidth}@{\quad}p{0.45\linewidth}@{}}
  \subfigimg[width=\linewidth]{(a)}{figs/shear_adR1em1_ra090_phaseDiagram.pdf} &
  \subfigimg[width=\linewidth]{(b)}{figs/shear_adR4em1_ra090_phaseDiagram.pdf}
  \end{tabular}
  \caption{\label{fig:sflow_phase_diagram} The phase diagram of vesicle
  hydrodynamics in a shear flow with $\delta = 0.1$ (a) and $\delta =
  0.4$ (b) with reduced area $\Delta A = 0.9$. The critical Hamaker
  constant $\mathcal{H}$ for binding of two vesicles in a shear flow
  depends on the shear rate $\dot\gamma$.  For a doublet to form, lower
  separation distances require a stronger Hamaker constants at a given
  shear rate.  The stability limit appears to be linear, indicating that
  there is potentially a stability criteria that depends on a
  dimensionless number involving the ratio of the Hamaker constant and
  shear rate.}
\end{figure}
